### 2017-02-01 paper1.tex
Add files via upload

Although most thermal remote sensors operate within one of the \gls{atmwind}[s] --- where atmospheric effects are greatly reduced --- virtually any remote sensed TIR signal is subject to radiative effects from the intervening atmosphere between the surface and the sensor. Over much of the thermal infrared \gls{waveband} the atmosphere emits radiation and absorbs a fraction of radiation emitted by the surface. In addition, the earth (and indeed the urban surface) reflects a portion of incoming solar radiation. Thus, a TIR signal measured from a remote sensing platform is almost certainly not equal to the ground emitted signal. These effects can lead to differences between the 'true' radiometric T\textsubscript{surf} and the remote sensed T\textsubscript{surf} of over 10 k for satellite platforms \cite{Cooper1989} and over 6 K for near-ground sensors \cite{Meier2011}. Moreover, because atmospheric and emissivity effects are a function of non-uniform and spatiotemporally variant surface and atmospheric properties, their associated errors change depending on instrument type, study location, and ambient conditions. As intersite and time-sequential analysis is a significant goal of most thermal remote sensed studies (urban or otherwise) these effects cannot be ignored.

Spectral transmission of longwave radiation through a given layer of atmosphere varies significantly based on wavelength, path geometry, and column absorber content (primarily H\textsubscript{2}0, C0\textsubscript{2}, and to a lesser extent 0\textsubscript{3}, N\textsubscript{2}0, CO, CH\textsubscript{4}, and 0\textsubscript{2} \cite{Miskolczi1993}). Variation in band-by-band longwave transmittance with path length is greatest at urban scales (approximately 1 to 50 meters), where transparent spectral bands can quickly become opaque with small changes in path length or absorber content - illustrated in figure \ref{spectransheight}. Thus, transmittance of TIR near the surface is highly dependent on surface-sensor-sun geometry, instrument properties, and synoptic conditions. Indeed, accurate assessment of atmospheric influence on TIR may be most complex when measured near the surface.

\begin{figure}[H]
	\includegraphics[width=\textwidth]{spectransheight}
	\caption{Spectral transmission of water vapor as a function of height. Black and grey shading indicate opaque (0\% transmittance) and transparent (100\% transmittance) bands respectively. Redo this figure with wavelength and colorbar?}
	\label{spectransheight}
\end{figure}

The radiation passing through a layer of atmosphere from an emitting surface can be described in a number of ways. \textit{Spectral directional radiance} $ R (\lambda)$, at height $ z $ and from a direction defined by zenith angle \(\theta\), and azimuth angle \(\phi\) is commonly written as

@@

A large number of correction routines have been developed to remove atmospheric and emissivity effects from aerial and satellite TIR signals and derive accurate T\textsubscript{rad}. Methods range from simple mono- \cite{Qin2001} and split-window \cite{Wan1996} routines for single- and multi-channel remote sensors to schemes that integrate a radiative transfer code to isolate the surface emitted signal from interfering signals. Boundary conditions are standard across most correction methods: often requiring vertical profiles of T\textsubscript{air}, humidity, pressure, and aerosol content to remove atmospheric effects, and assessments of surface radiative properties to correct for emissivity effects. However, correction methods are generally instrument (or at least platform) specific and difficult to generalize across sensor and platform types. Few methods exist for correction of ground based remote sensed TIR --- none of which are robust enough to correct irradiances upwelling from rough terrain measured  wide-\gls{fov}(FOV) sensor. In part, this is due to the fact that near ground thermal sensors have only recently become sensitive enough to require correction. However, a new breed of more accurate radiometers and thermal imagers should prompt a critical reevaluation of this assumption.

Atmospheric correction of near-ground remote sensed TIR is subject to a unique set of challenges compared to satellite and aerial platforms. Near-ground remote sensors have complex, multiple line-of-sight (LOS) geometry - illustrated in figure 2.1 for a wide-FOV instrument. Surface to sensor path lengths vary significantly over the sensor FOV, even when measured close (less than 5m) above the surface. This creates the potential for non-uniform atmospheric effects over the sensor FOV and necessitates a multi-LOS correction to retrieve accurate T\textsubscript{rad}. The addition of 3-dimensional surface geometry further amplifies this effect, as some paths may intersect with raised vertical, sloped, and horizontal features. In effect, with near-ground wide-FOV sensors, surface geometry is non-trivial and must be represented in atmospheric correction routines. In contrast, over a scene retrieved via satellite, spatial viability in surface geometry and LOS angle have a negligible effect on path length. Atmospheric correction routines for satellite retrieved TIR, therefore, assume uniform or single-LOS geometry because the TIR signal passes through a relatively constant volume of atmosphere over the projected sensor FOV, regardless of surface geometry. This greatly reduces the complexity of correction routines for satellite or aerial thermography.

@@

Radiative transfer of broadband longwave radiation in highly dependent on sensor-surface geometry. To calculate surface-sensor path lengths, the Surface-Sensor-Sun Urban Model (S\textsuperscript{3}UM) \cite{Soux2004} is initialized with a simplified, orthogonal 3-dimensional DBM representing the area immediately surrounding the sensor. In S\textsuperscript{3}UM, urban features (roads, roofs, buildings, etc.) are constructed from an array of horizontal and stacked vertical patches (representing rooftops/roads and walls respectively). After indicating sensor FOV and position over the DBM, the model determines which patches have an unobstructed line of sight to the sensor and calculates the path distance from "seen" patches to the sensor. Path lengths are binned at 5$^{\circ}$ increments from the nadir and averaged to return an azimuthally averaged path length for each 5$^{\circ}$ bin. Using azimuthally averaged path lengths in lieu of a full 3-dimensional radiative transfer model significantly reduces the computational time required to model each irradiance-T\textsubscript{hem} pairing, as at-sensor radiance can be computed independent of azimuth angle. Radiative transfer is thus reduced to a two dimensional problem (shown in panel (b) of figure \ref{23radtran}) in which an azimuthally averaged radiance is simulated for each zenith angle. The view factors required to weight directional radiances in the hemispherical integrated are computed While generating azimuthally averaged path lengths, S\textsuperscript{3}UM calculates view factors for each solid angle bin used in weighting 2-dimensional angular radiances during the hemispherical integration.  By using azimuthally averaged path lengths in lieu of  However, because hemispherical TIR sensor returns a bulk irradiance, the sensor signal does not need to be computed for each point of the

@@

\subsection{Modeling hemispherical irradiances}
With path length geometries calculates inS\textsuperscript{3}UM, MODTRAN 4.1 is initialized with 30-minute averaged profiles of T\textsubscript{air} and humidity to model at-sensor spectral directional radiances over a waveband of 0---2500cm\textsuperscript{-1}. Aerosol and above canyon meterological conditions are informed by the Mid-Latitude Summer standard atmosphere when dayttime maximum T\textsubscript{air} is greater than 10$^{\circ}$. For days with dayttime T\textsubscript{air} maximum below 10$^{\circ}$, the Mid-Latitude Winter standard atmosphere is substituted. In spite of significant computational efficiency gains from reducing 3-dimensional radiative transfer to azimuthally averaged 2-dimensions, the method still requires significant computational time to write input cards and model spectral radiances - in this case, with a path length increment of 5 and a LUT of 31 modeled irradiance---T\textsubscript{hem} pairings, each time-step requires 496 independent MODTRAN runs

>>>>

Although most thermal remote sensors operate within one of the \gls{atmwind}[s] --- where atmospheric effects are greatly reduced --- virtually any remote sensed TIR signal is subject to radiative effects from the layer of atmosphere between the surface and the sensor. Over much of the thermal infrared \gls{waveband} the atmosphere emits radiation and absorbs a fraction of radiation emitted by the surface. Thus, a remote sensed TIR signal almost certainly not equal to the ground emitted signal. For an isothermal, homogeneous surface-atmosphere system, at-sensor spectral radiance at height ($z$) can be described by a function deviating from a Planck curve at T\textsubscript{surf} based on the spectral transmittance of the intervening atmosphere, with the magnitude of that deviation governed by the difference between Planck curves at T\textsubscript{surf} and the ambient T\textsubscript{air}. As shown for two path geometries in \ref{spectairtsurf} for an atmosphere with water vapor content of  12 g/m\textsuperscript{3} and aerosol and trace gas profiles from the mid-latitude summer standard atmosphere, an at-sensor spectral radiance signal deviates significantly from the surface emitted spectral radiance curve. A less transmissive atmosphere increases the potential for deviation in the at-sensor signal from the ground emitted signal, while the difference between T\textsubscript{air} and T\textsubscript{surf} determines the resulting magnitude of atmospheric influence on the spectral TIR signal.

\begin{figure}[!ht]
	\includegraphics[width=\textwidth]{spectairtsurf}
	\caption{At-sensor spectral directional radiances computed in MODTRAN 4.1 \cite{Berk1987} for short (30m) and long (3000m) path lengths (z) with Planck curves indicating spectral radiances at T\textsubscript{air} = 305 K and T\textsubscript{surf} = 295 K.}
	\label{spectairtsurf}
\end{figure}

Atmospheric effects can lead to differences between the 'true' radiometric T\textsubscript{surf} and the remote sensed T\textsubscript{surf} of over 10 k for satellite platforms \cite{Cooper1989} and over 6 K for near-ground sensors \cite{Meier2011}. Moreover, because atmospheric and emissivity effects are a function of non-uniform and spatiotemporally variant surface and atmospheric properties, their associated errors change depending on instrument type, surface-sensor geometry, study location, and ambient conditions. As intersite and time-sequential analysis is a significant goal of most thermal remote sensed studies (urban or otherwise) these effects cannot be ignored.

Spectral transmission of longwave radiation through a given layer of atmosphere is dependent on total column absorber content (the principal broadband TIR absorbers are H\textsubscript{2}0, C0\textsubscript{2}, and to a lesser extent 0\textsubscript{3}, N\textsubscript{2}0, CO, CH\textsubscript{4}, and 0\textsubscript{2} \cite{Miskolczi1993}). Holding vertical absorber content constant, variation in band-by-band TIR transmittance with path length is greatest at urban scales (approximately 1 to 50 meters), where transparent spectral bands can quickly become opaque with small changes in path length or absorber content - illustrated in figure \ref{spectransheight}. Thus, transmittance of TIR near the surface is highly dependent on surface-sensor geometry, instrument spectral response, and atmospheric absorber content. Indeed, accurate assessment of atmospheric influence on TIR may be most complex when measured near the surface.

\begin{figure}[H]
	\includegraphics[width=\textwidth]{spectransheight}
	\caption{Spectral transmission of water vapor as a function of height. Black and grey shading indicate opaque (0\% transmittance) and transparent (100\% transmittance) bands respectively. Redo this figure with wavelength and colorbar no totle and the tick marks in back on the outside?}
	\label{spectransheight}
\end{figure}

Radiation passing through a layer of atmosphere from an emitting surface can be described in a number of ways. \textit{Spectral directional radiance} $ R (\lambda)$, at height $ z $ and from a direction defined by zenith angle \(\theta\), and azimuth angle \(\phi\) is commonly written as

@@

A large number of correction routines have been developed to remove atmospheric and emissivity effects from aerial and satellite TIR signals and derive accurate T\textsubscript{rad}. Methods range from simple mono- \cite{Qin2001} and split-window \cite{Wan1996} routines for single- and multi-channel remote sensors to schemes that integrate a radiative transfer code to isolate the surface emitted signal from interfering signals. Boundary conditions are standard across most correction methods: generally requiring vertical profiles of T\textsubscript{air}, humidity, pressure, and aerosol content to remove atmospheric effects, and assessments of surface radiative properties to correct for emissivity effects. However, correction methods are often instrument (or at least platform) specific and difficult to generalize across sensor and platform types. Few methods exist for correction of ground based remote sensed TIR --- none of which are robust enough to correct irradiances upwelling from rough terrain measured via wide-\gls{fov} (FOV) radiometers. In part, this is due to the fact that until recently, errors inherent in radiometer measurements were large relative to atmospheric effects. However, a new breed of more accurate radiometers and thermal imagers should prompt a critical reevaluation of this assumption.

Atmospheric correction of near-ground remote sensed TIR is subject to a unique set of challenges compared to traditional satellite and aerial platforms. Near-ground and wide-FOV remote sensors have complex, multiple line-of-sight (LOS) geometry - illustrated in figure 2.1 for a downward facing hemispherical radiometer. Surface-sensor geometry varies significantly over the sensor FOV, even when measured close (less than 5m) above the surface. The addition of 3-dimensional surface geometry further amplifies this effect, as some paths may intersect with raised vertical, sloped, and horizontal features. This creates the potential for non-uniform atmospheric effects over the sensor FOV and necessitates a multi-LOS correction to retrieve accurate T\textsubscript{rad}. In effect, with near-ground wide-FOV sensors, surface geometry is non-trivial and must be represented in atmospheric correction routines. In contrast, over a scene retrieved via satellite, spatial viability in surface geometry and LOS angle have a negligible effect on path length. Atmospheric correction routines for satellite retrieved TIR, therefore, assume uniform or single-LOS geometry because the TIR signal passes through a relatively constant volume of atmosphere over the projected sensor FOV, regardless of surface geometry. This greatly increases the complexity of correction routines for near-ground wide-FOV radiometry.

@@

The sheer number of unique path length geometries inherent with wide-FOV radiometry of rough terrain makes full 3-dimensional radiative transfer simulation difficult and computationally intensive --- particularly when correcting a climatology of T\textsubscript{hem}. In this method, to improve model efficiency, radiances are calculated for azimuthally averaged path lengths that represent average surface terrain for each solid angle "slice" of the sensor FOV. Thus, hemispherical radiative transfer is reduced to a 2-dimensional problem (shown in panel (b) of figure \ref{23radtran}). This greatly reduces the computational time required to model each irradiance-T\textsubscript{hem} pairing, as angular radiances can be computed as a function of zenith angle alone and subsequently weighted and integrated 3-dimensionally over the hemisphere.

To calculate surface-sensor geometries, the Surface-Sensor-Sun Urban Model (SUM) \cite{Soux2004} is initialized with a simplified, orthogonal 3-dimensional DBM representing the street canyons and courtyards immediately surrounding the sensor. SUM uses a four-dimensional array to represent surface morphology   with three spatial dimensions ($x$, $y$, and $z$), with $z$ representing height above the $x$, $y$ plane. SUM also includes an additional fourth dimension, where properties describing each cell are stored (in this case, distance from the point to the sensor). After specifying sensor FOV and position relative to the DBM, the model determines which patches have an unobstructed line of sight to the sensor and calculates the distance from "seen" patches to the sensor. Path lengths are binned at 5$^{\circ}$ increments of zenith angle and averaged to return an azimuthally-independent mean path length for each bin. In addition, while simulating path length geometries, SUM calculates view factors for each solid angle "slice", which are later used to weight 2-dimensional angular radiances in the hemispherical integration post-processing steps.

@@

\subsection{Modeling hemispherical irradiances}
With path length geometries calculated in SUM, irradiances are modeled time-sequentially using the MODerate resolution atmospheric TRANsmission 4.1 radiative transfer code (MODTRAN) \cite{Berk1987}. At each time-step, a range of potential T\textsubscript{hem} is selected using T\textsubscript{bright} subtracted by some constant (6 K during nighttime runs and 4 K for daytime runs). In MODTRAN, at-sensor spectral radiances for each path length are modeled over a waveband of 0---2500cm\textsuperscript{-1} at 0.5 K intervals over the range of potential T\textsubscript{hem}. Atmospheric profiles are constructed from 30-minute averages of T\textsubscript{air} and humidity with aerosol and above-sensor conditions are informed by the mid-latitude summer standard atmosphere when daytime maximum T\textsubscript{air} is greater than 10 $^{\circ}$C (the mid-latitude winter profile is substituted on days with T\textsubscript{air, max} of less than 10 $^{\circ}$C). 
Spectral radiances are convolved by a dome transmittance.
